---
title: "STA 160 Project"
author: "Jessica Padolina"
date: "4/30/2017"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}
setwd("/Users/Jessica/Documents/Stats Stuff/STA-160")
la_crime = read.csv("la_crime.csv")
```

CLEANING DATA
```{r}
#Subsetting the data to contain violent crimes only

la_violent = subset(la_crime, la_crime$Crime.Code.Desc %in% 
                     c("ASSAULT WITH DEADLY WEAPON ON POLICE OFFICER","ASSAULT WITH                         DEADLY WEAPON, AGGRAVATED ASSAULT", "ASSAULT", "BATTERY -                           SIMPLE ASSAULT", "BATTERY ON A FIREFIGHTER",
                       "BATTERY POLICE (SIMPLE)", "BATTERY WITH SEXUAL CONTACT",                           "BATTERY", "CRIMINAL HOMICIDE",
                       "MANSLAUGHTER, NEGLIGENT", "HOMICIDE", "ROBBERY"))

#Create pattern for Assault, Battery, and Homicide
#Robberies already correspond to one Crime.Code.Desc string, so no pattern necessary
pathA = "ASSAULT WITH DEADLY WEAPON ON POLICE OFFICER|ASSAULT WITH DEADLY WEAPON, AGGRAVATED ASSAULT" # Assault pathway

pathB = "BATTERY - SIMPLE ASSAULT|BATTERY ON A FIREFIGHTER|BATTERY POLICE \\(SIMPLE\\)|BATTERY WITH SEXUAL CONTACT" # Battery Pathway"
  
pathH = "CRIMINAL HOMICIDE|MANSLAUGHTER, NEGLIGENT" # Homicide Pathway" 
  
# Now use gsub() to replace each pattern with standardized crimet type
la_violent$Crime.Code.Desc = gsub(pathA, "ASSAULT", la_violent$Crime.Code.Desc) # sub for Assault

la_violent$Crime.Code.Desc = gsub(pathB, "BATTERY", la_violent$Crime.Code.Desc) # sub for Battery

la_violent$Crime.Code.Desc = gsub(pathH, "HOMICIDE", la_violent$Crime.Code.Desc) # sub for Homicide

#Splitting Coordinates column
library(stringr)
latlon = str_split_fixed(la_violent$Location.1, " ", 2)
latlon[,1] = gsub("\\(", "", latlon[,1])
latlon[,1] = gsub(",", "", latlon[,1])
latlon[,2] = gsub("\\)", "", latlon[,2])

#Converting to numeric class
latlon = apply(latlon, 2, as.numeric)

#Creating Year column
Year = substr(la_violent$Date.Occurred, 7, 10)
la_violent["Year"] <- Year
```

PLOTS
```{r}
#TABLE OF COUNTS OF CRIME PER TYPE
table(la_violent$Crime.Code.Desc)

#TIME PLOT OF ALL VIOLENT CRIME OVER TIME
#Group by Month and Year and create a new column
Mon = substr(la_violent$Date.Occurred, 1,2)
Year = substr(la_violent$Date.Occurred, 7, 10)

MonYr = paste(Mon, Year, sep = "/")
la_violent["MonYr"] <- MonYr

VCount = table(la_violent$MonYr)
VCount = data.frame(VCount)

#Change column names
colnames(VCount) = c("Date", "Count")

# Change the class of the variable 
VCount$Date = as.character(VCount$Date)

# Create a function to create a new date column
# original Date varaible is month/year
# we want to change it month/1 ( day)/ year
# reason: way to aggregate the data

monthly = function(data_col) {
  split_dates = strsplit(data_col, "/") # will split the month/year
  new_dates = sapply(split_dates, function(x) { 
    new_date = paste0(x[2], "-", x[1], "-01") # this adds the new date 
    new_date # now we have our new date = month/01/year
  })
  as.Date(new_dates)
}

VCount$Date = monthly(as.character(VCount$Date)) # run the Monthly function, variable must be a character

# order the dates
VCount = VCount[order(VCount$Date), ]

# Create the Time Series plot
TimeSeriesPlot = 
  plot_ly(VCount, x = VCount$Date, y = VCount$Count, 
          type = "scatter", mode = "lines", name = "Crime Count", line = list(color = 'rgba(67,67,67,1)', width = 2)) %>%
  layout(title = "Violent Crime 2010-2017",
         xaxis = list(title = "Date"), 
         yaxis = list(title = "Count")
  )

TimeSeriesPlot

#BAR CHARTS SHOWING COUNTS PER AREA
#Subset the data by type of Violent Crime
Assault = subset(la_violent, la_violent$Crime.Code.Desc ==  "ASSAULT")
Battery = subset(la_violent, la_violent$Crime.Code.Desc ==  "BATTERY")
Homicide = subset(la_violent, la_violent$Crime.Code.Desc ==  "HOMICIDE")
Robbery = subset(la_violent, la_violent$Crime.Code.Desc ==  "ROBBERY")


# Step 1) 
#  * Table to Dataframe Function * # 

vType = function(VTypeName, dataVar1, dataVar2){
  # create table by year
  VTypeName = table(dataVar1, dataVar2)   
  
  # create dataset for the table
  VTypeName = data.frame(VTypeName)
  
  # change column names
  colnames(VTypeName) = c("Area", "Year", "Count") 
  
  ## Pivot the table 
  Pivot = spread(VTypeName, Year, Count) 
  Pivot
}


# Step 2
# Now create the pivoted data frames for each violent crime
BPivot = vType(BType, Battery$Area.Name , Battery$Year)   # Battery
APivot = vType(AType, Assault$Area.Name , Assault$Year)   # Assault
HPivot = vType(HType, Homicide$Area.Name, Homicide$Year)  # Homicide
RPivot = vType(RType, Robbery$Area.Name , Robbery$Year)   # Robbery 


#Step 3
#  * Barchart Function * # 
# The function will later be used for the 4 diff violent crimes

ChartFunc =  
  # The function will read in each pivoted df
  # input each Var1 etc represent each var per year in the df
  function(chartName, titleName, data, dataVar1, dataVar2,
           dataVar3, dataVar4, dataVar5, dataVar6,
           dataVar7, dataVar8, dataVar9) { 
    
    chartName = 
      
      # use Plot_ly functions
      plot_ly(data, x = dataVar1) %>%
      add_bars(y = dataVar2, name = "2010") %>% # allows you to continue and start new line 
      add_bars(y = dataVar3, name = "2011", visible = FALSE) %>% # add each bar for each year
      add_bars(y = dataVar4, name = "2012", visible = FALSE) %>%
      add_bars(y = dataVar5, name = "2013", visible = FALSE) %>%
      add_bars(y = dataVar6, name = "2014", visible = FALSE) %>%
      add_bars(y = dataVar7, name = "2015", visible = FALSE) %>%
      add_bars(y = dataVar8, name = "2016", visible = FALSE) %>%
      add_bars(y = dataVar9, name = "2017", visible = FALSE) %>%
  
      layout(
        # Set up the layout 
        title = titleName, 
        xaxis = list(domain = c(0.1, 1)),
        yaxis = list(title = "Frequency"),
        updatemenus = list(
          list(
            y = 0.7,
            
            # Set up the drop down button per year
            # this will allow the user to control which year to view at a time
            
            buttons = list(
              list(method = "restyle",
                   args = list("visible", list(TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE)),
                   
                   label = "2010"),
              list(method = "restyle",
                   args = list("visible", list(FALSE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE)),
                   label = "2011"),
              list(method = "restyle",
                   args = list("visible", list(FALSE, FALSE, TRUE, FALSE, FALSE, FALSE, FALSE, FALSE)),
                   label = "2012"),
              list(method = "restyle",
                   args = list("visible", list(FALSE, FALSE, FALSE, TRUE, FALSE, FALSE, FALSE, FALSE)),
                   label = "2013"),
              list(method = "restyle",
                   args = list("visible", list(FALSE, FALSE, FALSE, FALSE, TRUE, FALSE, FALSE, FALSE)),
                   label = "2014"),
              list(method = "restyle",
                   args = list("visible", list(FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE, FALSE)),
                   label = "2015"),
              list(method = "restyle",
                   args = list("visible", list(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE, FALSE)),
                   label = "2016"),
              list(method = "restyle",
                   args = list("visible", list(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, TRUE)),
                   label = "2017")))
        )
      ) 
    chartName
  }

# Step 4
# Output for the barchart per Violent Crime

AssaultChart = 
  ChartFunc(AssaultChart2, "Crimes per Community Area",  
            APivot, APivot$Area, APivot$`2010`, APivot$`2011`,
            APivot$`2012`, APivot$`2013`, APivot$`2014`, 
            APivot$`2015`, APivot$`2016`, APivot$`2017`)

BatteryChart = 
  ChartFunc(BatteryChart2, "Crimes per Community Area",
            BPivot, BPivot$Area, BPivot$`2010`, BPivot$`2011`,
            BPivot$`2012`, BPivot$`2013`, BPivot$`2014`, 
            BPivot$`2015`, BPivot$`2016`, BPivot$`2017`)

HomicideChart = 
  ChartFunc(HomicideChart2, "Crimes per Community Area",
            HPivot, HPivot$Area, HPivot$`2010`, HPivot$`2011`,
            HPivot$`2012`, HPivot$`2013`, HPivot$`2014`, 
            HPivot$`2015`, HPivot$`2016`, HPivot$`2017`)

RobberyChart = 
  ChartFunc(RobberyChart2, "Crimes per Community Area",
            RPivot, RPivot$Area, RPivot$`2010`, RPivot$`2011`,
            RPivot$`2012`, RPivot$`2013`, RPivot$`2014`,
            RPivot$`2015`, RPivot$`2016`, RPivot$`2017`)

p <- subplot(AssaultChart, BatteryChart, HomicideChart, RobberyChart)
p


#TIME SERIES PLOT BROKEN DOWN BY TYPE OF CRIME
VType = table(la_violent$Crime.Code.Desc, la_violent$MonYr)
VType = data.frame(VType)

# Change the column names
colnames(VType) = c("Type", "Date", "Count")

# Change the class of the variable 
VType$Date = as.character(VType$Date)

VType$Date = monthly(as.character(VType$Date)) # run the Monthly function, variable must be a character

# Tidy R  pivot the data frame
VPivot = spread(VType, Type, Count)

# order the dates
VPivot = VPivot[order(VPivot$Date), ]

# Create the Time Series plot
TimeSeriesPlot2 = 
  plot_ly(VPivot, x = VPivot$Date, y = VPivot$ASSAULT, 
          type = "scatter", mode = "lines", name = "Assault", line = list(color = 'rgba(67,67,67,1)', width = 2)) %>%
  # add the remaining response variables variables to the plot
  add_trace(y = VPivot$BATTERY, name = "Battery", modes = "lines", 
            line = list(color = "dark blue"), width = 2 ) %>%
  add_trace(y = VPivot$HOMICIDE, name = "Homicide", modes = "lines", 
            line = list(color = "purple")) %>%
  add_trace(y = VPivot$ROBBERY, name = "Robbery", modes = "lines", 
            line = list(color = "blue")) %>%
  layout(title = "Violent Crime by Type 2010-2017",
         xaxis = list(title = "Date"), 
         yaxis = list(title = "Count")
  )

TimeSeriesPlot2
```

PREPARING DATA FOR MAPPING
```{r}
#Splitting the Coordinates column into two lat and lon columns (character matrix)
library(stringr)
latlon = str_split_fixed(la_violent$Coordinates, " ", 2)
latlon[,1] = gsub("\\(", "", latlon[,1])
latlon[,1] = gsub(",", "", latlon[,1])
latlon[,2] = gsub("\\)", "", latlon[,2])

#Converting to numeric class
latlon = apply(latlon, 2, as.numeric)

#Binding new latlon matrix to la_violent dataframe
la_violent[,c("lat","lon")] <- latlon

attach(la_violent)

#Deleting outliers (too far east of main incidences)
la_violent = subset(la_violent, lon < -118.1)

attach(la_violent)

#Creating Year column
Year = substr(Date.Occurred, 7, 10)
la_violent["Year"] <- Year
attach(la_violent)
```

DATA EXPLORATION
Preparing tract boundaries
```{r}
#MAPPING LA CRIME
#See STA 141A Lecture 11 slides
#Loading packages
library(ggmap)
library(rgdal)
library(scales)
library(dplyr)
library(Cairo)

#Preparing boundary shapefile

#Help from: https://stackoverflow.com/questions/29872109/binning-longitude-latitude-labeled-data-by-census-block-id (Code "Version 1")
tract <- readOGR(dsn = ".", layer = "census2010")
tract <- spTransform(x=tract, CRSobj=CRS("+proj=longlat +datum=WGS84"))
#tract <- fortify(tract, region = "GEOID10")

#Subsetting crimes by year
la_2010 = subset(la_violent, Year == "2010")
la_2011 = subset(la_violent, Year == "2011")
la_2012 = subset(la_violent, Year == "2012")
la_2013 = subset(la_violent, Year == "2013")
la_2014 = subset(la_violent, Year == "2014")
la_2015 = subset(la_violent, Year == "2015")
la_2016 = subset(la_violent, Year == "2016")
la_2017 = subset(la_violent, Year == "2017")

#Subsetting data

lonlat <- function(x) {
  y = data.frame(lat = x$lat, lon = x$lon)
  return(y)
}

lonlat.df10 <- lonlat(la_2010)
lonlat.df11 <- lonlat(la_2011)
lonlat.df12 <- lonlat(la_2012)
lonlat.df13 <- lonlat(la_2013)
lonlat.df14 <- lonlat(la_2014)
lonlat.df15 <- lonlat(la_2015)
lonlat.df16 <- lonlat(la_2016)
lonlat.df17 <- lonlat(la_2017)

#Converting data into a spatial points object
spdf <- function(x,y) {
  z = SpatialPointsDataFrame(coords=x[, c("lon", "lat")],
                                  data=y, proj4string=CRS("+proj=longlat +datum=WGS84"))
  return(z)
}

la_spdf10 <- spdf(lonlat.df10, la_2010)
la_spdf11 <- spdf(lonlat.df11, la_2011)
la_spdf12 <- spdf(lonlat.df12, la_2012)
la_spdf13 <- spdf(lonlat.df13, la_2013)
la_spdf14 <- spdf(lonlat.df14, la_2014)
la_spdf15 <- spdf(lonlat.df15, la_2015)
la_spdf16 <- spdf(lonlat.df16, la_2016)
la_spdf17 <- spdf(lonlat.df17, la_2017)

#Adding count column (each row is one crime incident)
spdf.list = list(la_spdf10,la_spdf11,la_spdf12,la_spdf13,la_spdf14,la_spdf15,la_spdf16,la_spdf17)

spdf <- function(x) {
  x@data$count <- 1
  return(x)
}

spdf.list <- lapply(spdf.list, spdf)

#Spatial overlay to identify census polygon in which each crime point falls
la_spdf_tract10 <- over(x=spdf.list[[1]], y=tract)
la_spdf_tract11 <- over(x=spdf.list[[2]], y=tract)
la_spdf_tract12 <- over(x=spdf.list[[3]], y=tract)
la_spdf_tract13 <- over(x=spdf.list[[4]], y=tract)
la_spdf_tract14 <- over(x=spdf.list[[5]], y=tract)
la_spdf_tract15 <- over(x=spdf.list[[6]], y=tract)
la_spdf_tract16 <- over(x=spdf.list[[7]], y=tract)
la_spdf_tract17 <- over(x=spdf.list[[8]], y=tract)

#Reassigning/delisting for ease of use
la_spdf10 = spdf.list[[1]]
la_spdf11 = spdf.list[[2]]
la_spdf12 = spdf.list[[3]]
la_spdf13 = spdf.list[[4]]
la_spdf14 = spdf.list[[5]]
la_spdf15 = spdf.list[[6]]
la_spdf16 = spdf.list[[7]]
la_spdf17 = spdf.list[[8]]

#Adding tract data to crime points
la_spdf10@data <- data.frame(la_spdf10@data, la_spdf_tract10)
la_spdf11@data <- data.frame(la_spdf11@data, la_spdf_tract11)
la_spdf12@data <- data.frame(la_spdf12@data, la_spdf_tract12)
la_spdf13@data <- data.frame(la_spdf13@data, la_spdf_tract13)
la_spdf14@data <- data.frame(la_spdf14@data, la_spdf_tract14)
la_spdf15@data <- data.frame(la_spdf15@data, la_spdf_tract15)
la_spdf16@data <- data.frame(la_spdf16@data, la_spdf_tract16)
la_spdf17@data <- data.frame(la_spdf17@data, la_spdf_tract17)

#Aggregate crimes by tract
agg = function(x) {
  y = aggregate(formula=count~GEOID10+Year, data=x@data, FUN=length)
  return(y)
}

ct10 <- agg(la_spdf10)
ct11 <- agg(la_spdf11)
ct12 <- agg(la_spdf12)
ct13 <- agg(la_spdf13)
ct14 <- agg(la_spdf14)
ct15 <- agg(la_spdf15)
ct16 <- agg(la_spdf16)
ct17 <- agg(la_spdf17)

#Merging all elements in ct.list
ct.list <- list(ct10,ct11,ct12,ct13,ct14,ct15,ct16,ct17)
ct = ct.list[[1]]
for (i in 2:length(ct.list)) {
  ct = rbind(ct, ct.list[[i]])
  i = i+1
}

#Sorting ct by Tract ID, then by Year (descending)
ct = sqldf("SELECT * FROM ct ORDER BY GEOID10, Year DESC;")
```

Trying various animated choropleth tutorials
```{r}
#Help from: https://gist.github.com/LaDilettante/5855730
rm(list=ls())
toInstall <- c("ggplot2", "maptools", "animation", "rgdal")
lapply(toInstall, library, character.only=TRUE)

#Load tract shape file
tract <- fortify(tract, region = "GEOID10")

# Plot

saveSWF({ # Output a flash
  for (i in c(2010:2017)) {
    temp <- subset(na.omit(la_violent[, c('DR.Number','Crime_Code','','year', 'country', 'cc_est')]), year==i)
    choro <- merge(world.ggmap, temp, by.x="id", by.y="country")
    choro <- choro[order(choro$order), ] # order matters in map data, so we must reorder
    p <- ggplot(data=choro, aes(long, lat, fill=cc_est, group=group))
    print(
      p + geom_polygon() + ylim(c(-60,85)) + # Fix the y-axis range for consistency between the years
        scale_fill_gradient2("Control of\n corruption\n score", limits=c(-2.5,2.5)) + # Use scale_fill_gradient2 for contrast
        labs(title=substitute(paste("Severity of corruption in the world, year ", i), list(i=i)))
    )
  }
}, swf.name="world_corruption.swf", interval=1,
        ani.width=1200, ani.height=600,
        swftools="C:/Program Files (x86)/SWFTools",
         outdir=getwd())

saveLatex({ # Output a pdf that can be embedded in latex documents
  for (i in c(1996, 1998, 2000, 2002:2011)) {
    temp <- subset(na.omit(wgi[, c('year', 'country', 'cc_est')]), year==i)
    choro <- merge(world.ggmap, temp, by.x="id", by.y="country")
    choro <- choro[order(choro$order), ]
    p <- ggplot(data=choro, aes(long, lat, fill=cc_est, group=group))
    print(
      p + geom_polygon() + ylim(c(-60,85)) +
        scale_fill_gradient2("Control of\n corruption\n score", limits=c(-2.5,2.5)) + 
        labs(title=substitute(paste("Severity of corruption in the world, year ", i), list(i=i)))
    )
  }
}, ani.basename = "world_corruption", interval=1,
          ani.width=900, ani.height=450,
          ani.opts="autoplay,loop,controls,width=\\linewidth",
          latex.filename = "world.corruption.tex",
outdir=getwd())

```

```{r}
#Help from: http://rmaps.github.io/blog/posts/animated-choropleths/
#Discretizing crime counts. We will assign 5 quantiles to the continuous values of "count" (counts of crimes)
ct2 <- transform(ct,
  fillKey = cut(count, quantile(count, seq(0, 1, 1/5)), labels = LETTERS[1:5])
)

#Associating fill colors
fills = setNames(
  c(RColorBrewer::brewer.pal(5, 'YlOrRd'), 'white'),
  c(LETTERS[1:5], 'defaultFill')
)

#Creating payload for DataMaps
#According to help source: "The data frame needs to be converted into a list of lists, as it is the default data structure that the DataMaps library accepts."

library(plyr)
require(devtools)
install_github('ramnathv/rCharts@dev')
install_github('ramnathv/rMaps')
library(rMaps)
library(rCharts)

dat2 <- dlply(na.omit(ct2), "Year", function(x){
  y = toJSONArray2(x, json = F)
  names(y) = lapply(y, '[[', 'GEOID10')
  return(y)
})

#Visualizing as a simple chloropleth
options(rcharts.cdn = TRUE)
map10 <- Datamaps$new()
map10$set(
  dom = 'chart_1',
  scope = 'usa',
  fills = fills,
  data = dat2[[1]],
  legend = TRUE,
  labels = TRUE
)
map10
```

Separate mapping step
```{r}
#Help source: http://www.kevjohnson.org/making-maps-in-r/
#Joining crime_tract and tract on tract id's
colnames(tract)[6] = "GEOID10"
plotdata = left_join(tract, crime_tract)

#Plotting for 2010
p <- ggplot() +
    geom_polygon(data = plotdata, aes(x = long, y = lat, group = group,
        fill = count), color = "black", size = 0.25)
ggsave(p, file = "map1.png", width = 6, height = 4.5, type = "cairo-png")
```

Plotting LA census tracts on map object
```{r}
library(maptools)
library(RColorBrewer)
CArea = readShapePoly("census2010")
CArea <- spTransform(x=tract, CRSobj=CRS("+proj=longlat +datum=WGS84"))
area.points <- fortify(CArea)
la = get_map("Los Angeles, CA")
ggmap(la)
colors <- brewer.pal(9, "BuGn")
ggmap(la) + geom_polygon(aes(x = long,
                   y = lat,
                   group = group),
               data = area.points,
               color = colors[9],
               fill = colors[6],
               alpha = 0.5) +
  labs(x = "Longitude",
       y = "Latitude")
```

Animating counts of crimes per tract over time
```{r}
#0. Finish making crime_tract dataframes
#1. Add date column to each crime_tract dataframe
#2. Combine all crime_tract dataframes into one yearly_counts dataframe

```

#Possible next steps:
#Center maps
#Include boundaries, multiple variables per map
#Subset by type of crime
#Do for remaining years
#Color by gender/race of victim (ditch area id)
#Animated plot rather than multiple plots per panel
#Consider distorting map to account for population density (cartogram)
#Finalize exploratory data analysis for LA
#See what patterns you can recognize, continue analysis using that

#Further subsetting by types of violent crime
#Types: 

#To include on animated plot:


```
